\bf Nonparametric Bootstrap Kolmogorov-Smirnov Goodness-of-Fit Test for Marginal Distributions of Stationary Time Series

### 2024-03-15 manuscript/manuscript.tex
little bit to conclusion

If we can show that the test correctly fails to reject $H_0$ under the null
distribution, and correctly rejects $H_0$ for an alternate distribution, we can
state that the method works.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$, Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
distributions.
We generated time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming the $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(X_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the unit exponential 
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.607, 
diverges, -0.287, 0, 0.339, 0.668, 0.911\}$.
\mc{my explanation needs work}

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is normal, and when $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is gamma. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

Like we did to see if the test holds its size, we generated time series with 
marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, Kendall's $\tau \in \{-.75, -.5, -.25, 0, .25, .5, 
.75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. However, 
to evaluate the test's power, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution family is gamma. Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0.
When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is normal. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

@@

\subsection{Precipitation at Airports}
\label{sec:precipitation}
Annual maximums of hourly precipitation in inches time series data were 
retrieved 
from the Chicago Midway International Airport (MDW), New York LaGuardia 
International 
Airport (LGA),
and Los Angeles International Airport (LAX) stations from the 
National Oceanic and Atmospheric Administration (NOAA). These stations were selected 
from places of importance in the three largest American cities. LGA and
MDW were chosen over John F. Kennedy and O'Hare as more data was available
from these stations. The data for MDW is from
1948 to 2014, the data for LGA is from 1948 to 2013, and the
data for LAX is also from 1948 to 2013. We can apply our method (as well
as the non-parametric bootstrap KS Test \citep{babu2004goodness} and
the parametric bootstrap KS Test \citep{zeimbekakis2022misuses}) to
the precipitation time series from each of these locations to test if they
follow the
Generalized Extreme Value (GEV) distribution using the \textsl{evd} 
package \citep{evd}. Precipitation maximum series 
typically do follow the GEV distribution, so we expect to fail to reject the
null hypothesis.

\jy{Is the parametric method by Zeimberkakis?}
\mc{yes}

%\input{tables/precipitation_pvals}
%\mc{https://hdsc.nws.noaa.gov/pfds/pfds_map_cont.html?bkmrk=ct}

The p-value is 0.71 for MDW, 0.13 for LGA, and 0.36 for LAX when using our 
method.
When using Babu's non-parametric method which does not account for serial
dependence, the p-value is 0.74 for MDW, 0.19 for LGA, and 0.40 for LAX. 
When using the parametric bootstrap KS test, the p-value is 0.64 for MDW,
0.12 for LGA, and 0.35 for LAX. In this instance, the conclusions are the
same 
(the three series follow the GEV distribution), but because of the difference
in p-values, it is possible for different conclusions to be found in another
study.

\subsection{Temperature in Central Park}
\label{sec:temperature}

Annual temperature maximums recorded in Central Park, New York, NY from
1895 to 2023 were sourced from the NOAA. Again, we are interested in seeing
if our method, which
accounts for serial dependence and does not require parameter specification,
performs differently than the non-parametric bootstrap KS Test 
\citep{babu2004goodness} and the parametric bootstrap KS Test, neither of
which account for serial dependence. We found that when using our method,
the p-value for the test that the series follows the GEV distribution is 0.1420.
When using Babu's method, the p-value is 0.0666, and when using the parametric
method, the p-value is 0.0682. At the common significance level of .05,
the conclusions would be the same regardless of the method used. However, it
is clear that for this series, our method more decisively fails to reject the 
null hypothesis at a significance level of .10.

Annual temperature minimums were also recorded from Central Park, and the
same method comparison was applied to this series. We found that our method
yields a p-value of 0.0015, Babu's method yields a p-value of 0.2747, and
the parametric method yields a p-value of 0.0002.

\subsection{Temperature in Eastern Asia}
\label{sec:asia}

Annual maximums of daily maximum (TXx) and minimum (TNx) temperatures, and annual 
minimums of daily maximum (TXn) and minimum (TNn) temperatures were collected
for 60 years from 44 locations in Eastern Asia.
\jy{Sorry I forgot to send you the data I promised. Remind in the future
  please. }

\subsection{Microsoft Stock Returns}
\label{sec:microsoft}

%\mc{https://www.nasdaq.com/market-activity/stocks/msft/historical}

We expect the marginal distribution of Microsoft stock returns to have a 
slightly
heavier tail than the normal distribution. Using the \textsl{tseries} package 
\citep{tseries}, 
we gathered closing stock prices from January 1st, 1998 to December 31st, 2022.
Then we approximated stock returns by taking the difference in the logarithm 
of the closing prices.
We tested that the stock returns are normally
distributed (which we expect to reject).  After that, we tested that the
marginal distribution of the time series follows the Student's $t$ distribution
with a fixed degrees of freedom.
While we 
could choose to estimate a fitted value of the degrees of freedom parameter $v$
as
part of our method, it is very difficult to do so. Therefore, we attempted to
test 
if it follows
the location-scale $t$ distribution with increasingly smaller specified degrees 
of freedoms. This was achieved with the \textsl{extraDistr} 
package \citep{extraDistr}. We
also observed how the method performs using increasingly smaller durations of
time. ,
we used the last 5 years of data: January 1st, 2018 to December 31st, 2022,
4 years of data: January 1st, 2019 to December 31st, 2022,
3 years of data: January 1st, 2020 to December 31st, 2022,
2 years of data: January 1st, 2021 to December 31st, 2022,
and 1 year of data: January 1st, 2022
to December 31st, 2022. In addition, we compared it to results from 
the Babu non-parametric and 
the parametric bootstrap KS Test.

\input{tables/mc_pval_1y}

\input{tables/mc_pval_2y}

\input{tables/mc_pval_3y}

\input{tables/mc_pval_4y}

\input{tables/mc_pval_5y}

\jy{Report the autocorrelation estimate of residual and residual squared in the
  text.}

The p-values of our test when applied to Microsoft stock return data using
different settings are showcased in 
Tables~\ref{table:microsoft1}--\ref{table:microsoft2}. Although the p-values
are indeed different between the three methods, the conclusions at a .05 
significance 
level are mostly the same. This motivated an additional application on SP500
and SP100 data.

@@

Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. It requires a larger sample size 
than basic bootstrap to avoid either type I or type II errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust. We have also 
demonstrated possible applications of this method to climate and financial data.
Although in many cases, the conclusions may not differ between our method
and other bootstrap KS tests, the p-values are different enough where in a case
where the temporal dependence in the data is more extreme, conclusions could
be drastically different if a different method is used.
This method can be used for any study where the goal is
to see if a time series follows a hypothesized distribution and the parameters
are unknown or unspecified. Future studies could apply the method showcased
in this paper to fields like earthquake prediction and astronomy.

>>>>

In this simulation study, we aim to show that the test correctly fails to 
reject $H_0$ under the null
distribution, and correctly rejects $H_0$ for an alternate distribution. If
both of these things are true, we can state that the method works
reasonably well.

@@

We generated time series with marginal distributions $N(8, 8)$ and
$\Gamma(8, 1)$ with Kendall's
\jy{Need to state how serial dependence were introduced, like you did in the
  first paper: transforming a normal series.}
$\tau \in \{-.75, -.5, -.25, 0, .25, .5, 75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. Kendall's $\tau$ was chosen as a
measure of serial dependence as it does not vary between two different 
marginal distributions.
To generate the samples to which our
method would be applied, we simulated a time series $W_t$ from a 1st 
order autoregressive (AR(1)) process:
\begin{equation*}
W_t = \phi X_{t-1} + \epsilon_t,
\end{equation*}
where $\phi$ is an autoregressive coefficient, and $\epsilon_t$ is a series of
independent errors from a normal distribution with mean zero and variance
$\sigma_{\epsilon}^2$. The strength of the serial dependence is controlled by
$\phi$, which was set to five levels: 
$\{-0.924, -0.707, -0.383, 0, 0.383, 0.707, 0.924\}$, as these
correspond to the desired values for $\tau$. First, we generated a
marginal $N(8, 8)$ by marginally transforming $W_t$ by
\begin{equation*}
X_t = F^{-1}[\Phi(W_t)],
\end{equation*}
where $F^{-1}(p)$ is the quantile function for the $N(8, 8)$ 
distribution.
Then we generated a marginal gamma series by the same procedure, but
replacing $F^{-1}(p)$ with the quantile function for the $\Gamma(8, 1)$
distribution.
After the transformation
to the marginal gamma distribution, the autocorrelations are $\{-0.607, 
diverges, -0.287, 0, 0.339, 0.668, 0.911\}$.
\mc{my explanation needs work}

These distributions
were chosen to compare results on normal and non-normal
error structures. The specific parameters were chosen because the distributions 
are very similar and their
first two moments are the same. 
For the purposes of 
evaluating if the test holds it size, when $X_t \sim N(8, 8)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. When $X_t \sim \Gamma(8, 1)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$.
For the block
bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method 1000 times 
to get the distribution of the p-values 
for the test when applied to samples from the same data generating process.

@@

Like we did to see if the test holds its size, we generated time series with 
marginal distributions $N(8, 8)$ and 
$\Gamma(8, 1)$, Kendall's $\tau \in \{-.75, -.5, -.25, 0, .25, .5, 
.75\}$, and
sample size $n \in \{100, 200, 400, 800\}$. However, 
to evaluate the test's power, when $X_t \sim N(8, 8)$, we tested
that the marginal distribution family is from the Gamma family,
or
$X_t \sim \Gamma(\cdot \mid \alpha, \beta), \quad t = 1, \ldots, n$.
for some unspecified $\alpha$ and $\beta$. Because the support of $N(8, 8)$ is
$(-\infty, \infty)$, but the support of $\Gamma(8, 1)$ is $(0, \infty)$, we used
the \textsl{truncdist} package \citep{truncdist} to truncate the series at 
values less than or equal to 0.
When $X_t \sim \Gamma(8, 1)$, we tested that the 
marginal distribution is from the
Normal family, or 
$X_t \sim N(\cdot \mid \mu, \sigma^2), \quad t = 1, \ldots, n$
for some unspecified $\mu$ and $\sigma$. For the block bootstrap step,
we used $B = 1000$ and $l = \lceil n^{1/3} \rceil$.
For each setting for $F$, $\tau$, and $n$, we replicate the method $1000$ times 
to obtain a p-value $p_r$ for each $r \in \{1, \ldots, 1000\}$.

@@

@@

The Kolmogorov-Smirnov test provides a way for researchers in a broad array
of applications to compare the distributions of populations. The
method proposed in this paper offers a variant of the test that can
be utilized for serially dependent data without specifying parameters.
Using simulation, we have shown that given a large enough sample of a time 
series, the block 
bootstrap KS test can appropriately fail to reject the null hypothesis when the
series follows the null distribution. In addition, when the marginal
distribution is not
the one hypothesized, the test is powerful. It requires a larger sample size 
than basic bootstrap to avoid either type I or type II errors. Unsurprisingly,
the method performs better as the temporal dependence gets weaker. Through
simulation studies with Normal and Gamma-distributed data, we have shown that
that the method is robust. We have also 
demonstrated possible applications of this method to financial data.
Through comparison with \citep{babu2004goodness}'s non-parametric
bootstrap KS test and \citet{zeimbekakis2022misuses}'s parametric bootstrap
KS test, we have shown that our method can reach different conclusions which
would perhaps be missed if serial dependence was neglected and another
method was chosen.
This method can be used for any study where the goal is
to see if a serially dependenttime series follows a hypothesized distribution and the parameters
are unknown or unspecified. Future studies could investigate
the method's performance on different marginal distributions and
different dependence structures. One could also apply the method showcased
in this paper to fields like earthquake prediction and astronomy.
